Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit

### 2024-08-10 jso.tex
Matérn simulation figures
added Matérn simulation figures in R and R^2 (Figure 1 and Figure 2)

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\theoremstyle{plain}
\newtheorem{assumption}{\protect\assumptionname}
\newtheorem{claim}{\protect\claimname}
\newtheorem{condition}{\protect\conditionname}
\newtheorem{conjecture}{\protect\conjecturename}
\newtheorem{cor}{\protect\corollaryname}
\newtheorem{defn}{\protect\definitionname}
\newtheorem{example}{\protect\examplename}
\newtheorem{lem}{\protect\lemmaname}
\newtheorem{notation}{\protect\notationname}
\newtheorem{problem}{\protect\problemname}
\newtheorem{prop}{\protect\propositionname}
\newtheorem{rem}{\protect\remarkname}
\newtheorem{thm}{\protect\theoremname}
\providecommand{\assumptionname}{Assumption}
\providecommand{\claimname}{Claim}
\providecommand{\conditionname}{Condition}
\providecommand{\conjecturename}{Conjecture}
\providecommand{\corollaryname}{Corollary}
\providecommand{\definitionname}{Definition}
\providecommand{\examplename}{Example}
\providecommand{\lemmaname}{Lemma}
\providecommand{\notationname}{Notation}
\providecommand{\problemname}{Problem}
\providecommand{\propositionname}{Proposition}
\providecommand{\remarkname}{Remark}
\providecommand{\theoremname}{Theorem}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal of Multivariate Analysis}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num-names}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit},
  pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
  pdfkeywords={projection pursuit, jellyfish search optimiser
(JSO), optimisation, grand tour, high-dimensional data, exploratory data
analysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    projection pursuit \sep jellyfish search optimiser
(JSO) \sep optimisation \sep grand tour \sep high-dimensional data \sep 
    exploratory data analysis
\end{keyword}
\end{frontmatter}

\section{Introduction}\label{introduction}

@@

\begin{equation}\phantomsection\label{eq-optimization}{
\underset{A \in \mathcal{A}}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I_d
}\end{equation}

@@

\begin{tcolorbox}[enhanced jigsaw, breakable, arc=.35mm, colframe=quarto-callout-note-color-frame, colback=white, rightrule=.15mm, coltitle=black, toprule=.15mm, toptitle=1mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, opacityback=0, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacitybacktitle=0.6, bottomtitle=1mm, bottomrule=.15mm, left=2mm]

@@

An classical way to describe the smoothness of a function is to identify
how many continuous derivatives of the function exist. This can be
characterized by Sobolev spaces \citep{adams2003sobolev}.

@@

The Matérn covariance function can be expressed analytically when
\(\nu\) is a half-integer, the most popular values in the literature
being \(\frac{1}{2}\), \(\frac{3}{2}\) and \(\frac{5}{2}\)
\citep{rasmussen2006gaussian}. The parameter \(\nu\), called
\emph{smoothness parameter}, controls the decay of the covariance
function. As such, it is an appropriate measure of smoothness of a
random field. For example, \citet{karvonen2023asymptotic} showed that if
a function \(f\) has a Sobolev index of \(k\), then the smoothness
parameter estimate \(\nu\) in \eqref{eq:matern} cannot be asymptotically
less than \(k\). See the survey \citet{porcu2024matern} for additional
results on the connection between the Matérn model and Sobolev spaces.
An interesting result is that the asymptotic case
\(\nu\rightarrow\infty\) coincides with the Gaussian kernel:
\(K_\infty(u)=\exp(-u^{2}/2)\).

@@

It is expected that these two values should be both high in the case of
high squintability (fast increase in \(f\) early on), and both low in
the case of low squintability (any substantial increase in \(f\) happens
very late, close to the optimal angle). This suggests that their product
(\textbf{?@eq-squintability}) should provide a sensible measure of
squintability. The multiplicative constant \(4\), which can be deemed
arbitrary, does not change the interpretation of the squintability
metric \(\varsigma\); it is here to adjust the range of values of
\(\varsigma\) and simplify the explicit formula for \(\varsigma\)
obtained later on.

@@

To compute the squintability metric (\textbf{?@eq-squintability}) in
practice, several approaches are possible. The first one is to propose a
parametric model for \(f\), and use it to obtain an explicit formula for
\(\varsigma\). Numerical experiments suggest a scaled sigmoid shape as
described below. Define

\begin{equation}\phantomsection\label{eq-logistic}{
\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-parametric}{
f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
}\end{equation}

which depends on three additional parameters, \(\theta_1\),
\(\theta_2\), and \(x_{\max}\), such that \(f(0)=\theta_1\) and
\(f(x_{\max})=\theta_4\). Under the parametric model
(\ref{eq-parametric}), the squintability metric
(\textbf{?@eq-squintability}) can be shown to be equal to

\begin{equation}\phantomsection\label{eq-squintability-parametric}{
\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
}\end{equation}

@@

Alternatively, one can estimate (\textbf{?@eq-squintability}) in a
nonparametric way, for example by fitting \(f\) using kernel regression,
then numerically estimate the angle at which \(-f'\) attains its highest
value.

\section{Simulation details}\label{sec-sim-deets}

@@

To compute squintability, 50 random bases are sampled and interpolated
to the optimal basis with a step size of 0.005. Index values and
projection distances are calculated for these interpolated bases and the
index values are averaged with a bin width of 0.005. A four-parameter
scaled logistic function is fitted to the index values against
projection distances, estimated by non-linear least squares. The
squintability measure is then calculated as
\ref{eq-squintability-parametric}.

@@

\caption{\label{fig-squintability}Illustration of steps to calculate
squintability. For a given projection pursuit problem defined by the
shape to find, data dimension and the index function, 1) sample random
bases given the orthonormality and projection distance contraint, 2)
interpolate the sampled bases to the optimal basis and calculate the
projection distance and the index value for each interpolated basis. 3)
bin the index values by projection distances to obtain the average index
value for each bin, 4) fit the scaled sigmoid function in Equation
\textbf{?@eq-squintability} to the binned index values against
projection distances using non-linear least square, 5) calculate the
squintability measure using Equation \ref{eq-squintability-parametric}
with parameters estimated from the model.}

\end{figure}%

\section{Results}\label{sec-sim-res}

@@

\end{figure}%

The index properties, including smoothness and squintability, offer
numerical metrics to characterise the complexity of projection pursuit
optimisation problems. Table~\ref{tbl-smoothness-squintability} presents
the parameters for calculating both metrics estimated from the Gaussian
process (variance, range, smooth, and nugget) and the scaled logistic
function (\(\theta_1\) to \(\theta_4\)) for each case considered in the
simulation. The column ``smooth'' is used as the smoothness metric and
the column ``squint'' is calculated as equation
\ref{eq-squintability-parametric} as the squintability metric.
Table~\ref{tbl-mod-output} presents the results of fitting a generalised
linear model with a binomial family and a logit link function to a
sample of data in Table~\ref{tbl-mod-data}, where all the three
components (success rate, JS hyper-parameters, and index properties) are
combined. The model suggests that JSO success rate is positively
associated with the two hyper-parameters, as well as with the index
properties: smoothness and squintability. Specifically, using 10 more
jellyfish and 10 more tries increases the odd ratio of success by
24.11\% and 11.93\%, respectively. However, being flagged with long
runtime and an increase of data dimension reduce the success rate by
41.72\% and 53.36\%, respectively. The variable \texttt{squintability}
and \texttt{dimension} are significant, suggesting their importance
relative to JSO hyper-parameters in the optimisation success. In
defining a projection pursuit problem, factors such as the
shape-to-find, the index function used, and the data dimension,
determine the properties such as smoothness and squintability. These
metrics can then be compared across different problems to understand
their relative complexites. The results suggest that squintability has a
more significant influence than smoothness on the success rate of JSO.
Once the characteristics of the projection pursuit problem are fully
understood, increasing JSO hyper-parameters can enhance the search
effectiveness, however, it is also important to consider the resulting
increase in computational complexity.

\begin{table}

\caption{\label{tbl-smoothness-squintability}Parameters estimated from
the Gaussian process (including variance, range, smoothness, and nugget)
and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
pipe-finding and sine-wave finding problems. The column ``smooth'' and
``squint'' represent the smoothness and squintability measures.}

\centering{

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\theoremstyle{plain}
\newtheorem{assumption}{\protect\assumptionname}
\newtheorem{claim}{\protect\claimname}
\newtheorem{condition}{\protect\conditionname}
\newtheorem{conjecture}{\protect\conjecturename}
\newtheorem{cor}{\protect\corollaryname}
\newtheorem{defn}{\protect\definitionname}
\newtheorem{example}{\protect\examplename}
\newtheorem{lem}{\protect\lemmaname}
\newtheorem{notation}{\protect\notationname}
\newtheorem{problem}{\protect\problemname}
\newtheorem{prop}{\protect\propositionname}
\newtheorem{rem}{\protect\remarkname}
\newtheorem{thm}{\protect\theoremname}
\providecommand{\assumptionname}{Assumption}
\providecommand{\claimname}{Claim}
\providecommand{\conditionname}{Condition}
\providecommand{\conjecturename}{Conjecture}
\providecommand{\corollaryname}{Corollary}
\providecommand{\definitionname}{Definition}
\providecommand{\examplename}{Example}
\providecommand{\lemmaname}{Lemma}
\providecommand{\notationname}{Notation}
\providecommand{\problemname}{Problem}
\providecommand{\propositionname}{Proposition}
\providecommand{\remarkname}{Remark}
\providecommand{\theoremname}{Theorem}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal of Multivariate Analysis}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num-names}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit},
  pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
  pdfkeywords={projection pursuit, jellyfish search optimiser
(JSO), optimisation, grand tour, high-dimensional data, exploratory data
analysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    projection pursuit \sep jellyfish search optimiser
(JSO) \sep optimisation \sep grand tour \sep high-dimensional data \sep 
    exploratory data analysis
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\begin{equation}\protect\hypertarget{eq-optimization}{}{
\underset{A \in \mathcal{A}}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I_d
}\label{eq-optimization}\end{equation}

@@

\begin{tcolorbox}[enhanced jigsaw, left=2mm, toptitle=1mm, opacityback=0, opacitybacktitle=0.6, bottomrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, breakable, leftrule=.75mm, colback=white, rightrule=.15mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, coltitle=black, bottomtitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm]

@@

A classical way to describe the smoothness of a function is to identify
how many continuous derivatives of the function exist. This can be
characterized by Sobolev spaces \citep{adams2003sobolev}.

@@

The Matérn covariance function can be expressed analytically when
\(\nu\) is a half-integer, the most popular values in the literature
being \(\frac{1}{2}\), \(\frac{3}{2}\) and \(\frac{5}{2}\)
\citep{rasmussen2006gaussian}. The parameter \(\nu\), called
\emph{smoothness parameter}, controls the decay of the covariance
function. As such, it is an appropriate measure of smoothness of a
random field, as shown by the simulations on Figure~\ref{fig-matern-1d}
and Figure~\ref{fig-matern-2d}. For example,
\citet{karvonen2023asymptotic} showed that if a function \(f\) has a
Sobolev index of \(k\), then the smoothness parameter estimate \(\nu\)
in \eqref{eq:matern} cannot be asymptotically less than \(k\). See the
survey \citet{porcu2024matern} for additional results on the connection
between the Matérn model and Sobolev spaces. An interesting result is
that the asymptotic case \(\nu\rightarrow\infty\) coincides with the
Gaussian kernel: \(K_\infty(u)=\exp(-u^{2}/2)\).

{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_1d.png}

\caption{\label{fig-matern-1d}Five random simulations from a Gaussian
Process defined on \(\mathbb{R}\) with zero mean and Matérn-\(\nu\)
covariance function, with \(\nu=1\) (left), \(\nu=2\) (middle), and
\(\nu=4\) (right), showing that higher values of \(\nu\) produce
smoother curves.}

\end{figure}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_2d.png}

\caption{\label{fig-matern-2d}One random simulation from a Gaussian
Process defined on \(\mathbb{R}^2\) with zero mean and Matérn-\(\nu\)
covariance function, with \(\nu=1\) (left), \(\nu=2\) (middle), and
\(\nu=4\) (right), showing that higher values of \(\nu\) produce
smoother surfaces.}

\end{figure}

@@

It is expected that these two values should be both high in the case of
high squintability (fast increase in \(f\) early on), and both low in
the case of low squintability (any substantial increase in \(f\) happens
very late, close to the optimal angle). This suggests that their product
\eqref{eq-squintability} should provide a sensible measure of
squintability. The multiplicative constant \(4\), which can be deemed
arbitrary, does not change the interpretation of the squintability
metric \(\varsigma\); it is here to adjust the range of values of
\(\varsigma\) and simplify the explicit formula for \(\varsigma\)
obtained later on.

@@

To compute the squintability metric \eqref{eq-squintability} in
practice, several approaches are possible. The first one is to propose a
parametric model for \(f\), and use it to obtain an explicit formula for
\(\varsigma\). Numerical experiments suggest a scaled sigmoid shape as
described below. Define

\begin{equation}\protect\hypertarget{eq-logistic}{}{
\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
}\label{eq-logistic}\end{equation}

@@

\begin{equation}\protect\hypertarget{eq-parametric}{}{
f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
}\label{eq-parametric}\end{equation}

which depends on three additional parameters, \(\theta_1\),
\(\theta_2\), and \(x_{\max}\), such that \(f(0)=\theta_1\) and
\(f(x_{\max})=\theta_4\). Under the parametric model
(\ref{eq-parametric}), the squintability metric \eqref{eq-squintability}
can be shown to be equal to

\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
}\label{eq-squintability-parametric}\end{equation}

@@

Alternatively, one can estimate \eqref{eq-squintability} in a
nonparametric way, for example by fitting \(f\) using kernel regression,
then numerically estimate the angle at which \(-f'\) attains its highest
value.

\hypertarget{sec-sim-deets}{%
\section{Simulation details}\label{sec-sim-deets}}

@@

To compute squintability, 50 random bases are sampled and interpolated
to the optimal basis with a step size of 0.005. Index values and
projection distances are calculated for these interpolated bases and the
index values are averaged with a bin width of 0.005. A four-parameter
scaled logistic function is fitted to the index values against
projection distances, estimated by non-linear least squares. The
squintability measure is then calculated as
(\ref{eq-squintability-parametric}).

@@

\caption{\label{fig-squintability}Illustration of steps to calculate
squintability. For a given projection pursuit problem defined by the
shape to find, data dimension and the index function, 1) sample random
bases given the orthonormality and projection distance contraint, 2)
interpolate the sampled bases to the optimal basis and calculate the
projection distance and the index value for each interpolated basis. 3)
bin the index values by projection distances to obtain the average index
value for each bin, 4) fit the scaled sigmoid function in equation
\eqref{eq-squintability} to the binned index values against projection
distances using non-linear least square, 5) calculate the squintability
measure using equation (\ref{eq-squintability-parametric}) with
parameters estimated from the model.}

\end{figure}

\hypertarget{sec-sim-res}{%
\section{Results}\label{sec-sim-res}}

@@

\end{figure}

The index properties, including smoothness and squintability, offer
numerical metrics to characterise the complexity of projection pursuit
optimisation problems. Table~\ref{tbl-smoothness-squintability} presents
the parameters for calculating both metrics estimated from the Gaussian
process (variance, range, smooth, and nugget) and the scaled logistic
function (\(\theta_1\) to \(\theta_4\)) for each case considered in the
simulation. The column ``smooth'' is used as the smoothness metric and
the column ``squint'' is calculated as equation
(\ref{eq-squintability-parametric}) as the squintability metric.
Table~\ref{tbl-mod-output} presents the results of fitting a generalised
linear model with a binomial family and a logit link function to a
sample of data in Table~\ref{tbl-mod-data}, where all the three
components (success rate, JS hyper-parameters, and index properties) are
combined. The model suggests that JSO success rate is positively
associated with the two hyper-parameters, as well as with the index
properties: smoothness and squintability. Specifically, using 10 more
jellyfish and 10 more tries increases the odd ratio of success by
24.11\% and 11.93\%, respectively. However, being flagged with long
runtime and an increase of data dimension reduce the success rate by
41.72\% and 53.36\%, respectively. The variable \texttt{squintability}
and \texttt{dimension} are significant, suggesting their importance
relative to JSO hyper-parameters in the optimisation success. In
defining a projection pursuit problem, factors such as the
shape-to-find, the index function used, and the data dimension,
determine the properties such as smoothness and squintability. These
metrics can then be compared across different problems to understand
their relative complexites. The results suggest that squintability has a
more significant influence than smoothness on the success rate of JSO.
Once the characteristics of the projection pursuit problem are fully
understood, increasing JSO hyper-parameters can enhance the search
effectiveness, however, it is also important to consider the resulting
increase in computational complexity.

\hypertarget{tbl-smoothness-squintability}{}
\begin{table}
\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (including variance,
range, smoothness, and nugget) and scaled logistic function
(\(\theta_1\) to \(\theta_4\)) for the pipe-finding and sine-wave
finding problems. The column ``smooth'' and ``squint'' represent the
smoothness and squintability measures. }\tabularnewline
